The role of cognateness in non-native spoken word recognition

### 2022-10-28 manuscript.qmd
Add questionnaire data and analyses, and change reports to Quarto

shorttitle: Cognateness and non-native word recognition

author:
- name: Gonzalo Garcia-Castro
  affiliation: '1'
  corresponding: yes
  address: Ramon Trias Fargas, 25-27, 08005 Barcelona
  email: gonzalo.garciadecastro@upf.edu
- name: Serene Siow
  affiliation: '2'
- name: Nuria Sebastian-Galles
  affiliation: '1'
- name: Kim Plunkett
  affiliation: '2'
affiliation:
  - id            : "1"
    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
  - id            : "2"
    institution   : "Department of Experimental Psychology, University of Oxford"

authornote: |

@@

  <!-- https://tinyurl.com/ybremelq -->

keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
wordcount: 4553

bibliography:
  - r-references.bib
  - references.bib
floatsintext: yes
figurelist: no
tablelist: no
footnotelist: no
linenumbers: yes
mask: no
draft: no
numbersections: yes
csl: "apa7.csl"
documentclass: apa7
classoption: man
output: papaja::apa6_docx
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

set.seed(42)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  results = "asis", 
  dpi = 1000
)

@@

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# import objects
tar_load_globals()
tar_load_all()

# set custom ggplot2 theme
theme_set(
  theme(
    title = element_text(
      size = 8, 
      face = "bold"
    ),
    legend.background = element_rect(
      fill = NA,
      colour = NA
    ),
    legend.key = element_rect(
      fill = "white", 
      colour = NA),
    legend.justification = "right",
    legend.direction = "horizontal",
    axis.line = element_line(
      colour = "black",
      size = 0.5
    ),
    axis.ticks = element_line(
      colour = "black", 
      size = 0.5
    ),
    axis.text.x = element_text(
      colour = "black"
    ),
    axis.text.y = element_text(
      colour = "black"
    ),
    plot.caption.position = "plot",
    plot.title = element_text(
      hjust = 0.5
    ),
    panel.grid = element_blank(),
    panel.background = element_rect(
      fill = "white", 
      colour = NA
    )
  )
)

@@

Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`. We collected data from `r length(unique(participants$participant_id))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`). `r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant_id[participants$group %in% c("cat-SPA")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant_id[participants$group %in% c("cat-SPA") & participants$sex=="Female"]))` female). Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits). Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value. Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I). Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection. We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), (b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`), or (c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).

@@

```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}

include_graphics(here("img", "design.png"))

@@

```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lengths <- stimuli %>% 
  select(
    group, 
    ipa1,
    ipa2, 
    word1, 
    word2
  ) %>% 
  mutate_at(
    vars(ipa1:word2),
    nchar
  ) %>% 
  group_by(group) %>% 
  summarise_at(
    vars(
      ipa1, 
      ipa2, 
      word1, 
      word2
    ), 
    list(
      mean = mean, 
      sd = sd,
      min = min, 
      max = max
    )
  ) %>%
  split(.$group) %>% 
  map(select, -group) %>% 
  set_names(make_clean_names(names(.))) %>% 
  map(unlist)
```

@@

We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only). Words in the Catalan list were `r printnum(lengths$cat_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa1_max"], digits = 0)`). Words in the Spanish list were `r printnum(lengths$spa_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa1_max"], digits = 0)`). Participants listened to one audio file in each trial, each containing a single word presented in isolation. The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain). These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner. Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat]. The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`). The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`).

@@

Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word2_sd"])`, Range = `r printnum(lengths$cat_eng["word2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word2_sd"])`, Range = `r printnum(lengths$cat_spa["word2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word2_max"], digits = 0)`). Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word2_sd"])`, Range = `r printnum(lengths$spa_eng["word2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word2_max"], digits = 0)`).

@@

```{r pthn_example, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

pthn_example <- clearpond_path %>% 
  map(read.csv) %>% 
  bind_rows(.id = "group") %>% 
  clean_names() %>% 
  filter(
    word %in% c("book", "sock"),
    group=="cat-ENG"
  ) %>% 
  select(word, pthn)

@@

```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA). Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores. B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages. Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."}

d <- stimuli %>% 
  mutate(
    group = factor(
      group,
      levels = c(
        "cat-ENG",
        "spa-ENG", 
        "cat-SPA"
      )
    )
  )

@@

# frequency
d %>% 
  ggplot() +
  aes(
    frequency_zipf, 
    fill = group
  ) +
  facet_wrap(
    ~group,
    ncol = 1
  ) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = frequency_zipf_median-frequency_zipf_sd, 
      xmax = frequency_zipf_median+frequency_zipf_sd, 
      ymin = -Inf, 
      ymax = Inf, 
      fill = group
    ), 
    inherit.aes = FALSE,
    alpha = 0.15
  ) +   
  geom_histogram(
    bins = 20,
    colour = "white"
  ) +
  geom_vline(
    data = medians, 
    aes(
      xintercept = frequency_zipf_median,
      group = group
    ), 
    colour = "black",
    size = 0.5,
    linetype = "dashed"
  ) + 
  geom_text(
    data = medians, 
    aes(
      label = "Median ± SD", 
      x = frequency_zipf_median+1.25,
      y = 13
    ), 
    size = 3
  ) +
  geom_curve(
    data = medians, 
    aes(
      x = frequency_zipf_median+0.85, 
      xend = frequency_zipf_median+0.1,
      y = 12, 
      yend = 11
    ),
    arrow = arrow(length = unit(0.2, "cm")),
    curvature = -0.25,
    colour = "grey30"
  ) +
  labs(
    x = "Lexical frequency (Zipf)", 
    y = "Number of trials", 
    title = "Frequency\n(stimuli translations)\n"
  ) +

  # PTHN
  d %>% 
  count(
    group, 
    pthn
  ) %>% 
  ggplot() + 
  aes(
    pthn,
    n, 
    fill = group
  ) +
  facet_wrap(
    ~group,
    ncol = 1
  ) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = 0, 
      xmax = pthn_median+pthn_sd, 
      ymin = -Inf, 
      ymax = Inf,
      fill = group
    ), inherit.aes = FALSE, 
    alpha = 0.15
  ) +   
  geom_col(colour = "white") + 
  geom_vline(
    data = medians, 
    aes(
      xintercept = pthn_median, 
      group = group
    ), 
    colour = "black", 
    size = 0.5, 
    linetype = "dashed"
  ) + 
  labs(
    x = "PTHN", 
    y = "Number of trials", 
    title = "Higher-frequency neighbours\n(stimuli translations)\n"
  ) +

  # Levenshtein
  d %>% 
  ggplot() +
  aes(
    lv, 
    fill = group
  ) +
  facet_wrap(
    ~group,
    ncol = 1
  ) + 
  geom_rect(
    data = medians, 
    aes(
      xmin = 0,
      xmax = lv_median+lv_sd, 
      ymin = -Inf,
      ymax = Inf, 
      fill = group
    ),
    inherit.aes = FALSE,
    alpha = 0.15
  ) +   
  geom_histogram(
    bins = 20, 
    colour = "white"
  ) +
  geom_vline(
    data = medians, 
    aes(
      xintercept = lv_median,
      group = group
    ), 
    colour = "black",
    size = 0.5, 
    linetype = "dashed"
  ) + 
  scale_x_continuous(labels = percent) +
  labs(
    x = "Levenshtein", 
    y = "Number of trials", 
    title = "Phonological similarity\n(stimuli and stimuli translations)\n"
  ) +

  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "A") &
  scale_fill_manual(
    values = c("#1A85FF", "#ff2976", "#FFC20A"),
    na.translate = FALSE) &
  theme(
    legend.position = "none",
    title = element_text(size = 10),
    axis.title = element_text(size = 10)
  )
```

@@

```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

tab_loos <- model_loos %>% 
  as.data.frame() %>% 
  rownames_to_column("model") %>% 
  arrange(desc(elpd_loo)) %>% 
  select(
    model,
    matches("elpd_loo"),
    matches("looic"), 
    matches("diff")) %>%
  mutate(
    model = str_replace(
      model,
      "fit_", 
      "Model "
    )
  ) %>% 
  mutate_all(~ifelse(. == 0, NA, .)) %>% 
  gt(
    rowname_col = "model", 
    id = "model-comparison"
  ) %>% 
  tab_header(
    title = md("Table 2"),
    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
  ) %>% 
  tab_source_note(
    md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
  ) %>% 
  fmt_missing(
    everything(),
    missing_text = "-"
  ) %>% 
  fmt_number(2:7) %>% 
  cols_label(
    model = md("**Model**"),
    looic = md("***LOO<sub>IC</sub>***"),
    se_looic = md("***SE*<sub>IC</sub>**"),
    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
    se_elpd_loo = md("***SE***"),
    elpd_diff = md("***LOO<sub>diff</sub>***"),
    se_diff = md("***SE<sub>diff</sub>***")
  ) %>% 
  tab_style(
    style = cell_text(
      align = "left",
      weight = "bold", 
      size = "medium"
    ),
    locations = cells_title(
      groups = "title"
    )
  ) %>% 
  tab_style(
    style = cell_text(
      align = "left", 
      size = "medium", 
      style = "italic"
    ),
    locations = cells_title(
      groups = "subtitle"
    )
  ) %>% 
  tab_style(
    style = cell_text(
      align = "left", 
      size = "medium"
    ),
    locations = cells_source_notes()
  ) %>% 
  tab_style(
    style = list(
      cell_borders(
        sides = c("left", "right"),
        color = "white"
      ),
      cell_text(
        align = "center"
      )
    ),
    locations = list(
      cells_body(), 
      cells_row_groups(), 
      cells_column_spanners(), 
      cells_stub(), 
      cells_source_notes(),
      cells_column_labels(), 
      cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(
      cell_text(
        align = "left",
        size = "medium"
      ), 
      cell_borders(
        sides = "all", 
        color = "white"
      )
    ),
    locations = list(
      cells_stub(), 
      cells_source_notes()
    )
  ) %>% 
  tab_style(
    style = cell_text(
      align = "center",
      size = "medium", 
      weight = "normal",
      style = "normal"
    ),
    locations = list(
      cells_column_spanners(), 
      cells_column_labels()
    )
  ) %>% 
  tab_style(
    style = cell_borders(
      sides = "all",
      color = "white"
    ),
    locations = cells_body(
      columns = 1:7
    )
  ) %>% 
  tab_style(
    style = cell_borders(
      sides = "top", 
      color = "white"
    ),
    locations = cells_body(
      columns = 1:7
    )
  ) %>% 
  gtsave(here("img", "loos.png"))

@@

coefs <- fixef(fit_6) %>% 
  as.data.frame() %>% 
  rownames_to_column("variable") %>% 
  clean_names() %>% 
  mutate_at(
    vars(estimate:q97_5),
    ~ifelse(
      variable=="Intercept", 
      inv_logit_scaled(.), 
      ./4
    )
  ) %>% 
  group_split(variable) %>% 
  set_names(make_clean_names(map(., "variable"))) %>% 
  map(select, -variable) %>% 
  map(unlist)

@@

```{r bayesfactors, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# wrapper to extract Bayes Factors
get_bf <- function(hyp, abs = TRUE, ...){
  bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
  if (abs) bf <- abs(bf)
  return(bf)
}

@@

word_example <- ranef(fit_6)$translation_id[,,"Intercept"] %>% 
  as.data.frame() %>% 
  rownames_to_column("word") %>% 
  arrange(-Estimate) %>% 
  clean_names() %>% 
  mutate_if(is.numeric, inv_logit_scaled) %>% 
  slice(1, n()) %>% 
  left_join(
    distinct(
      stimuli,
      word1, 
      ipa1), 
    by = c("word" = "word1")) %>% 
  split(.$word)

@@

cors_range <- gather_draws(
  fit_6, 
  `cor_.*`, 
  regex = TRUE
) %>% 
  median_qi() %>% 
  summarise(.value = range(.value))

@@

\newpage

@@

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

\newpage

>>>>

author:
  - name: Gonzalo Garcia-Castro
    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
  - name: Serene Siow
    affiliation: Department of Experimental Psychology, University of Oxford
  - name: Nuria Sebastian-Galles
    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
  - name: Kim Plunkett
    affiliation: Department of Experimental Psychology, University of Oxford

@@

keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian

bibliography:
  - r-references.bib
  - references.bib
toc: false
number-sections: true
csl: "apa7.csl"
format:
  pdf:
    message: false
    warning: false
    results: "asis" 
    fig-dpi: 1000
---

```{r}
#| label: setup
#| message: false
#| warning: false
#| include: false

set.seed(42)

@@

```{r}
#| label: prepare
#| message: false
#| warning: false
# import objects
tar_load_globals()
tar_load_all()

# set custom ggplot2 theme
theme_set(
    theme(
        title = element_text(size = 8, face = "bold"),
        legend.background = element_rect(fill = NA, colour = NA),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.justification = "right",
        legend.direction = "horizontal",
        axis.line = element_line(colour = "black", size = 0.5),
        axis.ticks = element_line(colour = "black", size = 0.5),
        axis.text.x = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black"),
        plot.caption.position = "plot",
        plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white", colour = NA)
    )
)

@@

Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`. We collected data from `r length(unique(participants$participant_id))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`). `r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant_id[participants$group %in% c("cat-SPA")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant_id[participants$group %in% c("cat-SPA") & participants$sex=="Female"]))` female). Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits). Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value. Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I) and by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007). Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection. We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), (b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`), or (c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).

@@

```{r}
#| label: procedure-figure
#| echo: false
#| message: false
#| warning: false
#| out-width: "80%"
#| fig-cap: "Schematic representation of a trial in the experimental task."
include_graphics(here("img", "design.png"))
```

@@

```{r}
#| label: stimuli-lengths
#| message: false
#| warning: false
lengths <- stimuli %>% 
    select(group, ipa_flat_1, ipa_flat_2, word_1, word_2) %>% 
    mutate_at(vars(ipa_flat_1:word_2), nchar) %>% 
    group_by(group) %>% 
    summarise_at(
        vars(ipa_flat_1, ipa_flat_2, word_1, word_2), 
        list(mean = mean, sd = sd, min = min, max = max)
    ) %>%
    split(.$group) %>% 
    map(select, -group) %>% 
    set_names(make_clean_names(names(.))) %>% 
    map(unlist)
```

@@

We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only). Words in the Catalan list were `r printnum(lengths$cat_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa_flat_1_max"], digits = 0)`). Words in the Spanish list were `r printnum(lengths$spa_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa_flat_1_max"], digits = 0)`). Participants listened to one audio file in each trial, each containing a single word presented in isolation. The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain). These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner. Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat]. The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`). The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`).

@@

Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word_2_sd"])`, Range = `r printnum(lengths$cat_eng["word_2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word_2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word_2_sd"])`, Range = `r printnum(lengths$cat_spa["word_2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word_2_max"], digits = 0)`). Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word_2_sd"])`, Range = `r printnum(lengths$spa_eng["word_2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word_2_max"], digits = 0)`).

@@

```{r}
#| label: pthn-example
#| message: false
#| warning: false
pthn_example <- clearpond_path %>% 
    map(read.csv) %>% 
    bind_rows(.id = "group") %>% 
    clean_names() %>% 
    filter(word %in% c("book", "sock"), group=="cat-ENG") %>% 
    select(word, pthn)

@@

```{r}
#| label: stimuli-plot
#| message: false
#| warning: false
#| echo: false
#| out-width: "100%"
#| fig-width: 7
#| fig-height: 5.5
#| fig-cap: "Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA). Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores. B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages. Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."
d <-  mutate(stimuli, group = factor(group, levels = c("cat-ENG", "spa-ENG", "cat-SPA")))

@@

# frequency
d %>% 
    ggplot(aes(frequency_zipf, fill = group)) +
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(
            xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd,
            ymin = -Inf, ymax = Inf, fill = group
        ), 
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(
        bins = 20,
        colour = "white"
    ) +
    geom_vline(
        data = medians, 
        aes(xintercept = frequency_zipf_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    geom_text(
        data = medians, 
        aes(label = "Median ± SD", x = frequency_zipf_median+1.25, y = 13), 
        size = 3
    ) +
    geom_curve(
        data = medians, 
        aes(x = frequency_zipf_median+0.85, xend = frequency_zipf_median+0.1, y = 12, yend = 11),
        arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30"
    ) +
    labs(
        x = "Lexical frequency (Zipf)", 
        y = "Number of trials", 
        title = "Frequency\n(stimuli translations)\n"
    ) +

    # PTHN
    d %>% 
    count(group, pthn) %>% 
    ggplot(aes(pthn, n, fill = group)) + 
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(xmin = 0, xmax = pthn_median+pthn_sd, ymin = -Inf, ymax = Inf, fill = group),
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_col(colour = "white") + 
    geom_vline(
        data = medians, 
        aes(xintercept = pthn_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    labs(
        x = "PTHN", 
        y = "Number of trials", 
        title = "Higher-frequency neighbours\n(stimuli translations)\n"
    ) +

    # Levenshtein
    d %>% 
    ggplot(aes(lv, fill = group)) +
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(xmin = 0, xmax = lv_median+lv_sd, ymin = -Inf, ymax = Inf, fill = group),
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(bins = 20, colour = "white") +
    geom_vline(
        data = medians, 
        aes(xintercept = lv_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    scale_x_continuous(labels = percent) +
    labs(
        x = "Levenshtein", 
        y = "Number of trials", 
        title = "Phonological similarity\n(stimuli and stimuli translations)\n"
    ) +
    plot_layout(guides = "collect") &
    plot_annotation(tag_levels = "A") &
    scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
    theme(
        legend.position = "none",
        title = element_text(size = 10),
        axis.title = element_text(size = 10)
    )
```

@@

```{r}
#| label: model-comparison
#| message: false
#| warning: false
tab_loos <- model_loos %>% 
    as.data.frame() %>% 
    rownames_to_column("model") %>% 
    arrange(desc(elpd_loo)) %>% 
    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
    mutate_all(~ifelse(. == 0, NA, .)) %>% 
    gt(rowname_col = "model", id = "model-comparison") %>% 
    tab_header(
        title = md("Table 2"),
        subtitle = md("Model performance and comparison using leave-one-out cross-validation")
    ) %>% 
    tab_source_note(
        md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_number(2:7) %>% 
    cols_label(
        model = md("**Model**"),
        looic = md("***LOO<sub>IC</sub>***"),
        se_looic = md("***SE*<sub>IC</sub>**"),
        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
        se_elpd_loo = md("***SE***"),
        elpd_diff = md("***LOO<sub>diff</sub>***"),
        se_diff = md("***SE<sub>diff</sub>***")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = list(
            cell_borders(sides = c("left", "right"), color = "white"),
            cell_text(align = "center")
        ),
        locations = list(
            cells_body(), 
            cells_row_groups(), 
            cells_column_spanners(), 
            cells_stub(), 
            cells_source_notes(),
            cells_column_labels(), 
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"), 
            cell_borders(sides = "all", color = "white")
        ),
        locations = list(cells_stub(), cells_source_notes())
    ) %>% 
    tab_style(
        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
        locations = list(cells_column_spanners(), cells_column_labels())
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    gtsave(here("img", "loos.png"))

@@

coefs <- fixef(fit_6) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
    group_split(variable) %>% 
    set_names(make_clean_names(map(., "variable"))) %>% 
    map(select, -variable) %>% 
    map(unlist)

@@

```{r}
#| label: bayes-factors
#| message: false
#| warning: false
# wrapper to extract Bayes Factors
get_bf <- function(hyp, abs = TRUE, ...){
    bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
    if (abs) bf <- abs(bf)
    return(bf)
}

@@

word_example <- ranef(fit_6)$translation_id[,,"Intercept"] %>% 
    as.data.frame() %>% 
    rownames_to_column("word") %>% 
    arrange(-Estimate) %>% 
    clean_names() %>% 
    mutate_if(is.numeric, inv_logit_scaled) %>% 
    slice(1, n()) %>% 
    left_join(
        distinct(stimuli, word1, ipa1), 
        by = c("word" = "word1")
    ) %>% 
    split(.$word)

@@

cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
    median_qi() %>% 
    summarise(.value = range(.value))

@@

@@
